---
title: "Single cell marker prediction -- DREAM AIM I"
author: "Attila Gabor"
date: "4/26/2019"
output: html_document
---


```{r include=FALSE}
library(tidyverse)
library(DBI)
library(RSQLite)
library(progress)
```
Here we build random forest model to predict the single cell p-ERK response to combination of 
stimulus and inhibitors. Predictions are based on other markers. 
The model is trained from cell-lines which contains all the markerks in many exper


```{r}
con <- dbConnect(RSQLite::SQLite(), "./data/cleaned_single_cell_data/single_cell_dream_cls.sqlite")
```

Explore the database (tutorial-ish)

Get available cell-lines. Each cell-lines are stored in a table. 
```{r}
tables <-  dbListTables(con)
cols <- dbListFields(con,tables[[1]])
timepoints <- dbGetQuery(con,paste0("SELECT DISTINCT time FROM  ",dbQuoteIdentifier(con,tables[[1]])))
treatments <- dbGetQuery(con,paste0("SELECT DISTINCT treatment FROM  ",dbQuoteIdentifier(con,tables[[1]])))
```


## DREAM AIM 1 data preprocessing
First we load the cell-lines for the training and for the test.
We specify the features and response variables. 
Remove the response variables from the test set.

```{r}
training_celllines = c('AU565','BT483','CAL148', 'SKBR3','HBL100', 'HCC2185','MCF10A')
test_celllines = c('T47D','HCC1806','HCC2218', 'HCC70','UACC893')

training_size_per_cond = 500
test_size_per_cond = 500

bar = progress::progress_bar$new(format = "  Processing [:bar] :percent eta: :eta",
								 total = length(training_celllines))

training_data = sapply(training_celllines, function(cl){
	bar$tick()
	
	# load cell_line and limit the number of cells
	tmp = dbReadTable(con,  dbQuoteIdentifier(con,cl)) %>% 
		as_tibble() %>%
		group_by(treatment,time) %>% 
		filter(row_number()<training_size_per_cond)
	
	
},simplify = F)

training_data = do.call("rbind",training_data) %>% as_tibble()

bar = progress::progress_bar$new(format = "  Processing [:bar] :percent eta: :eta",
								 total = length(test_celllines))
test_data = sapply(test_celllines, function(cl){
	bar$tick()
	
	# load cell_line and limit the number of cells
	tmp = dbReadTable(con,  dbQuoteIdentifier(con,cl)) %>% 
		as_tibble() %>%
		group_by(treatment,time) %>% 
		filter(row_number()<test_size_per_cond)
},simplify = F)	

test_data = do.call("rbind",test_data) %>% as_tibble()


aim1_data = training_data %>% mutate(data_type = "training") %>%
	bind_rows(test_data %>% mutate(data_type = "test"))

```



```{r}
reporters <- colnames(training_data)[6:42]
#reporters <- setdiff(colnames(training_data)[6:42],c("p.HER2", "p.PLCg2"))
response = c("p.ERK", "p.Akt.Ser473.","p.S6","p.HER2", "p.PLCg2")
stopifnot(all(response %in% reporters))

predictors = setdiff(reporters, response)
```

## 1. Phospho-prediction on single cell level

Visualise the data outline as a table / heatmap
```{r}
pheat_data = aim1_data %>% 
	group_by(cell_line, treatment,data_type) %>% summarise_at(reporters,~mean(.)) %>% 
	ungroup() %>%
	gather("reporter","value",-cell_line,-treatment,-data_type) %>%
	mutate(condition = paste(reporter,treatment,sep = "_")) %>%
	select(reporter,cell_line,condition,value) %>%
	mutate(value = ifelse(reporter %in% response,-1,1)) %>%
	mutate(value = ifelse(cell_line %in% training_celllines,2,value)) %>%
	spread(cell_line,value,fill = NA) 

row_names = pheat_data %>% group_by(reporter) %>% mutate(row_name = ifelse(row_number() == round(n()/2),reporter,"")) %>%
	ungroup() %>%
	pull(row_name)

#column_to_rownames("row_name") %>%

pheat_res = pheat_data %>%
	select(-reporter,-condition) %>% 
	select(training_celllines,test_celllines)%>%
	pheatmap::pheatmap(mat = ., color = c("#E41A1C", "#377EB8", "#4DAF4A") ,
					   cluster_rows = F,
					   cluster_cols = F,
					   #   border_color=NA,
					   na_col = "grey", labels_row = row_names)
ggsave("./figures/aim1_data.pdf",plot = pheat_res$gtable,width = 5,height = 7)

```
Visualise with the emphasiz on single cell:

select a cell line from training and one from validation and show which data shall be imputed. 

```{r,fig.width=6, fig.height=3}
vis_data_1 <-  aim1_data %>%
	filter(cell_line %in% c("AU565","HCC1806")) %>%
	mutate(cell_line = ifelse(cell_line=="AU565","training","test")) %>%
	filter(time == 7, treatment=="iMEK") %>% ungroup %>%
	group_by(cell_line,time,treatment) %>% 
	filter(fileID == min(fileID)) %>% ungroup %>%
	filter(cellID < 100) %>%
	gather("reporter","value",-cell_line,-treatment,-time,-cellID,-fileID,-data_type) %>%
	filter(!  (data_type == "test" & reporter %in% response )) %>%
	filter(reporter %in%  c("p.ERK","p.S6","p.MEK","p.MKK4","p.p90RSK","p.S6K")) %>%
	mutate(fill_col = ifelse(reporter %in% c("p.ERK","p.S6"),"response","predictor")) %>%
	filter(!is.na(value))

gg_aim_1_scell_fig1 = vis_data_1  %>%	ggplot(data=., aes(factor(""),value)) +
	geom_dotplot(fill=rgb(colorRamp(colors=RColorBrewer::brewer.pal(11,"Spectral"))(vis_data_1$cellID/max(vis_data_1$cellID)),maxColorValue = 256),
				 binaxis='y',binwidth=0.1 , binpositions="all", stackdir='center',dotsize = 8,) + guides(fill="none")+
	facet_grid(cell_line ~ reporter) + theme_bw() + xlab("condition")
print(gg_aim_1_scell_fig1)
ggsave(plot=gg_aim_1_scell_fig1,filename = "./figures/aim1_single_cell_details_dotplot_v1.pdf",height = 3,width = 6)
```

For AIM 1.2 we cannot identify cells, so we can plot each cell without a color. 
we will compare distributions.


```{r,fig.width=6, fig.height=3}
vis_data_1  %>%	ggplot(data=., aes(factor(""),value)) +
	geom_dotplot(fill="white",#fill=rgb(colorRamp(colors=RColorBrewer::brewer.pal(11,"Spectral"))(vis_data_1$cellID/max(vis_data_1$cellID)),maxColorValue = 256),
				 #col=rgb(colorRamp(colors=RColorBrewer::brewer.pal(11,"Spectral"))(vis_data_1$cellID/max(vis_data_1$cellID)),maxColorValue = 256),
				 binaxis='y',binwidth=0.1 , binpositions="all", stackdir='center',dotsize = 5,) + guides(fill="none")+
	#geom_violin(aes(col=fill_col,fill=fill_col),alpha=.5,scale = "count",size=2) +
	facet_grid(cell_line ~ reporter) + theme_bw() + xlab("condition")

```


## 1.1 RF simple psites
First we use only the p-Sites (apart from p-ERK). 
We dont use explicitly the time or treatment in this model. 

### option 1: 
- predict p-ERK using all the samples, run all at once. 
```{r}
library(ranger)
# p-ERK
rangerres_res1=ranger(as.formula(paste(response[1], " ~ ", paste(predictors, collapse= "+"))),
					  data = aim1_data %>% filter(data_type =="training"), 
					  num.trees=300,
					  importance = "impurity",
					  verbose=T,
					  seed=12345,
					  num.threads = 3)
```

### option 2 (control): 
- predict p-ERK using all the samples, shuffled. Training data is column-wise shuffled, which 
destroys the correlation between markers and probably results in a useless model. 
```{r}
# p-ERK: shuffled training data
rand_shuffle <- function(x){
	sample(x,length(x))
}

randomised_data = aim1_data %>% filter(data_type =="training") %>%
	group_by(cell_line) %>% mutate_at(predictors,rand_shuffle) %>% ungroup()


# randomised_data %>% filter(time==7,treatment=="EGF", cell_line== "AU565") %>% ggplot() + geom_point(aes(p.ERK,p.p90RSK))
# aim1_data %>% filter(data_type =="training") %>% filter(time==7,treatment=="EGF", cell_line== "AU565") %>% ggplot() + geom_point(aes(p.ERK,p.p90RSK))

rangerres_res1_random=ranger(as.formula(paste(response[1], " ~ ", paste(predictors, collapse= "+"))),
							 data = randomised_data, 
							 num.trees=300,
							 importance = "impurity",
							 verbose=T,
							 seed=12345,
							 num.threads = 3)
rangerres_res1_random$prediction.error
rangerres_res1_random$r.squared
```


### option 3: 
p-ERK, but let's build a model for each condition:
```{r}
bar = progress::progress_bar$new(format = "  Processing [:bar] :percent eta: :eta",
								 total = nrow(unique(aim1_data %>% 
								 						filter(data_type =="training") %>%
								 						select("treatment","time"))
								 			 ))

fit_forest = function(tr_data){
	bar$tick()
	rf_model=ranger(as.formula(paste(response[1], " ~ ", paste(predictors, collapse= "+"))),
					data = tr_data,
					num.trees=300,
					importance = "impurity",
					verbose=T,
					seed=12345,
					num.threads = 3)
}


aim1_models_3 <- aim1_data %>% filter(data_type =="training") %>%
	group_by(treatment,time) %>% nest() %>%
	mutate(models = map(data,fit_forest))

aim1_models_3_stats <-  aim1_models_3 %>% mutate(model_stats  = map(models,function(m){
	as.data.frame(m[c("r.squared","prediction.error")])
} )) %>% unnest(model_stats) %>%
	mutate(std_y = map(data,function(d) std_y = sd(d$`p.ERK`))) %>%
	unnest(std_y) %>% select(treatment,time,r.squared, prediction.error,std_y)

aim1_models_3_stats %>% ggplot(aes(sqrt(1-r.squared)*std_y,prediction.error)) + geom_point()
aim1_models_3_stats %>% ggplot(aes(r.squared,prediction.error)) + geom_point()
```



```{r}

# p.Akt.Ser473.
rangerres_res2=ranger(as.formula(paste(response[2], " ~ ", paste(predictors, collapse= "+"))),
					  data = training_data, 
					  num.trees=300,
					  importance = "impurity",
					  verbose=T,
					  seed=12345,
					  num.threads = 3)

# p.S6
rangerres_res3=ranger(as.formula(paste(response[3], " ~ ", paste(predictors, collapse= "+"))),
					  data = training_data, 
					  num.trees=300,
					  importance = "impurity",
					  verbose=T,
					  seed=12345,
					  num.threads = 3)



```


This  shows that p-ERK is mainly determined by its downstream nodes


```{r}

pred_res1 <- predict(rangerres_res1, data = aim1_data %>% filter(data_type =="test") )
pred_res1_rand <- predict(rangerres_res1_random, data = aim1_data %>% filter(data_type =="test"))


#pred_res2 <- predict(rangerres_res2, data = validation_data)
#pred_res3 <- predict(rangerres_res3, data = validation_data)


comparison_data <-  aim1_data %>% filter(data_type =="test") %>% 
	select(cell_line,treatment,time,cellID,fileID,p.ERK) %>% #p.Akt.Ser473.,p.S6
	bind_cols(p.ERK.pred = pred_res1$predictions,
			  p.ERK.pred_random = pred_res1_rand$predictions) %>%
	gather("variable","value",starts_with("p.ERK")) %>%
	mutate(is_prediction = ifelse(grepl("pred",variable),"prediction","data")) %>%
	mutate(is_prediction = ifelse(grepl("pred_random",variable),"random",is_prediction))

#	



```

```{r, fig.height=6, fig.width=10}
# plot the p-ERK prediction and data
comparison_data %>% 
	#gather(key = "variable",value = "value",-cell_line,-treatment,-time,-cellID,-fileID) %>%
	filter(grepl("p.ERK",variable,fixed = T)) %>%
	filter(cell_line == "HCC70") %>%
	filter(treatment == "iPKC") %>%
	ggplot() +
	geom_violin(aes(x=as.factor(time), y=value,fill=is_prediction), scale = "count") +
	facet_grid(treatment~cell_line) +
	scale_fill_manual(values = c("data" = "black", "prediction" = "red","random"="green")) + 
	theme_bw() + xlab("time") + ylab("signal") + #guides(fill="none") + 
	ggtitle("p-ERK prediciton in HCC70, iPKC")
ggsave("./figures/aim1_HCC70_iPKC_timecourse.pdf",plot=gg,width = 7,height = 2)
```


```{r,fig.width=20}

comparison_data %>% 
	ggplot() +
	geom_violin(aes(x=1, y=value,fill=is_prediction), scale = "count") +
	facet_grid(treatment~cell_line+variable) +
	theme_bw() + theme(axis.text.x = element_blank(),axis.ticks.x =  element_blank()) + xlab("") + ylab("signal") + 
	ggtitle("RF prediction in test cell lines")
ggsave("./figures/aim1_RF_all_v1.pdf",plot = gg,width = 13,height = 6)
```

```{r, fig.height=30, fig.width=10}
# plot the p-ERK prediction and data
gg =comparison_data %>% 
	#filter(cell_line == "T47D") %>%
	ggplot() + 
	geom_point(aes(x=`p.ERK`, y=`p.ERK.pred`)) +
	facet_grid(cell_line+treatment~time) +
	theme_bw() + 
	ggtitle("p-ERK correlation in conditions")
```
# Metrics for AIM 1
A.  distance of predicted and measured markers in individual cells
B. correlation of predicted and measured markers
C. distance of distributions


## distance of predicted and measured markers in individual cells


```{r}
ucid = unique(comparison_data$cellID)
ggdat = comparison_data %>% 
	#gather(key = "variable",value = "value",-cell_line,-treatment,-time,-cellID,-fileID) %>%
	#mutate(is_prediction = grepl("pred",variable)) %>%
	filter(cellID <100) %>%
	filter(grepl("p.ERK",variable,fixed = T)) %>%
	filter(cell_line == "HCC70") %>%
	filter(treatment == "iPKC") %>%
	filter(time==0)

ggdat%>%ggplot() +
	geom_dotplot(aes(x=ifelse(is_prediction,"predicted","test"), y=value),
				 fill= rgb(colorRamp(colors=RColorBrewer::brewer.pal(11,"Spectral"))(ggdat$cellID/max(ggdat$cellID)),maxColorValue = 256),
				 binaxis='y',binpositions="all", stackdir='center',dotsize = 0.8) +
	theme_bw()  + xlab("") + ylab("signal") 

gg1 =  ggdat%>% filter(is_prediction) %>%ggplot() +
	geom_dotplot(aes(x=ifelse(is_prediction,"predicted","test"), y=value,fill=factor(cellID)),
				 stackgroups = TRUE,
				 binaxis='y',binpositions="all", stackdir='center',dotsize = 1.2) +
	theme_bw()  + xlab("") + ylab("signal") +guides(fill="none")+ ylim(0,5)
ggsave("./figures/aim1_HCC70_iPKT_T0_dot_pred.pdf",plot = gg1,width = 2.5,height = 4)

gg2 = ggdat%>% filter(!is_prediction) %>%ggplot() +
	geom_dotplot(aes(x=ifelse(is_prediction,"predicted","test"), y=value,fill=factor(cellID)),
				 stackgroups = TRUE,
				 binaxis='y',binpositions="all",stackdir='center',dotsize = 0.8) +
	theme_bw()  + xlab("") + ylab("signal") +guides(fill="none")+ ylim(0,5) 
ggsave("./figures/aim1_HCC70_iPKT_T0_dot_test.pdf",plot = gg2,width = 2.5,height = 4)

gg = ggdat%>%ggplot() +
	geom_jitter(aes(x=ifelse(is_prediction,"predicted","test"), y=value),width = 0.2,
				color= rgb(colorRamp(colors=RColorBrewer::brewer.pal(11,"Spectral"))(ggdat$cellID/max(ggdat$cellID)),maxColorValue = 256)) +
	theme_bw()  + xlab("") + ylab("signal") 


ggsave("./figures/aim1_HCC70_iPKT_T0_dot_v1.pdf",plot = gg,width = 5,height = 4)
```

compute RMSE for the this set:
```{r}
ggdat %>% select(-is_prediction) %>% spread(variable,value) %>% summarise(RMSE = sqrt(sum((p.ERK - p.ERK.pred)^2/n())))
```
```{r}
RMSE_data <- comparison_data %>% group_by(cell_line,time,treatment) %>%summarise(RMSE = sqrt(sum((p.ERK - p.ERK.pred)^2/n()))) %>% arrange(RMSE)
RMSE_data
```




## correlation of predicted and measured markers
```{r}
# plot the p-ERK prediction and data
corr_data = comparison_data %>% 
	filter(cellID <100) %>%
	filter(cell_line == "HCC70") %>%
	filter(treatment == "iPKC") %>%
	filter(time==0)
gg_corr <- corr_data %>%	ggplot(aes(x=`p.ERK`, y=`p.ERK.pred`)) +
	geom_point(col=rgb(colorRamp(colors=RColorBrewer::brewer.pal(11,"Spectral"))(corr_data$cellID/max(corr_data$cellID)),maxColorValue = 256)) +
	xlab("predicted") + ylab("test") + coord_equal() + xlim(0,6) + ylim(0,6) + geom_smooth(method="lm",fullrange=T) +
	theme_bw() 
ggsave("./figures/aim1_HCC70_iPKT_T0_correlation_v1.pdf",plot = gg_corr,width = 5,height = 4)
```

```{r}
corr_data %>% group_by(cell_line,time,treatment) %>%summarise(corr.coef = cor(p.ERK, p.ERK.pred)) %>% arrange(corr.coef)
corr_coeff_data<- comparison_data %>% group_by(cell_line,time,treatment) %>%summarise(corr.coef = cor(p.ERK, p.ERK.pred)) %>% arrange(corr.coef)
corr_coeff_data
```


```{r,fig.width=3,fig.height=3}
corr_coeff_data %>% full_join(RMSE_data, by = c("cell_line", "time", "treatment")) %>%
	ggplot(aes(corr.coef,RMSE)) + geom_point() + geom_smooth(method="lm") + theme_bw() + theme(aspect.ratio = 1)

RMSE_corr_stats=corr_coeff_data %>% full_join(RMSE_data, by = c("cell_line", "time", "treatment"))
cor.test(RMSE_corr_stats$RMSE,RMSE_corr_stats$corr.coef)
```


## Distance of distributions

- Kullbach-leibler distance
```{r}

#install.packages("philentropy")
library(philentropy)
library(purrr)
gg_dens <-  comparison_data %>% 
	filter(cellID <100) %>%
	filter(cell_line == "HCC70") %>%
	filter(treatment == "iPKC") %>%
	filter(time==0) %>% gather(variable,value,p.ERK,p.ERK.pred) %>%
	ggplot() + geom_density(aes(value,group=variable,fill=variable),alpha=0.5) + theme_bw()
ggsave("./figures/aim1_HCC70_iPKT_T0_densities_v1.pdf",plot = gg_dens,width = 5,height = 4)


KL_data = comparison_data %>% 
	filter(cellID <100) %>%
	filter(cell_line == "HCC70") %>%
	filter(treatment == "iPKC") %>%
	filter(time==0) %>% select(p.ERK,p.ERK.pred)

# estimate densitites at same points:
test_dens <- density(x = KL_data$p.ERK,from = 0,to = 10,n = 64)
pred_dens <- density(x = KL_data$p.ERK.pred,from = 0,to = 10,n = 64)
all(test_dens$x==pred_dens$x)

KL(rbind(test=test_dens$y,pred=pred_dens$y),est.prob = "empirical")
KL(rbind(pred=pred_dens$y,test=test_dens$y),est.prob = "empirical")
```




```{r}
library(purrr)
tmp <-  comparison_data %>% 
	select(-cellID,-fileID) %>%
	gather(reporter, value, -cell_line,-treatment,-time) %>%
	mutate(condition=ifelse(grepl("pred",reporter),"prediction","test")) %>%
	mutate(reporter = gsub(".pred","",reporter)) %>%
	group_by(cell_line,treatment,time,reporter) %>% nest() %>%
	mutate(density_data = map(data,function(df){
		
		test_data <- df %>% filter(condition=="test") %>% pull(value)
		prediction_data <- df %>% filter(condition=="prediction") %>% pull(value)
		test_dens <- density(x = test_data,from = 0,to = 10,n = 64)
		pred_dens <- density(x = prediction_data,from = 0,to = 10,n = 64)
		KL(rbind(test=test_dens$y,pred=pred_dens$y),est.prob = "empirical")
		
		
		
	}))



filter(cell_line == "HCC70") %>%
	filter(treatment == "iPKC") %>%
	filter(time==0) %>% gather(variable,value,p.ERK,p.ERK.pred) %>%
	ggplot() + geom_density(aes(value,group=variable,fill=variable),alpha=0.5) + theme_bw()
ggsave("./figures/aim1_HCC70_iPKT_T0_densities_v1.pdf",plot = gg_dens,width = 5,height = 4)


KL_data = comparison_data %>% 
	filter(cellID <100) %>%
	filter(cell_line == "HCC70") %>%
	filter(treatment == "iPKC") %>%
	filter(time==0) %>% select(p.ERK,p.ERK.pred)

# estimate densitites at same points:
test_dens <- density(x = KL_data$p.ERK,from = 0,to = 10,n = 64)
pred_dens <- density(x = KL_data$p.ERK.pred,from = 0,to = 10,n = 64)
all(test_dens$x==pred_dens$x)

KL(rbind(test=test_dens$y,pred=pred_dens$y),est.prob = "empirical")
KL(rbind(pred=pred_dens$y,test=test_dens$y),est.prob = "empirical")
```



## Correlation metric

```{r,fig.height=5,fig.width=5}
validation_data %>% filter(treatment=="EGF",time %in% c(0,9),cell_line == "HCC2218") %>%
	mutate(condition = ifelse(time==9,"test","prediciton")) %>% 
	ggplot(aes(p.ERK,p.p90RSK,group=condition)) + geom_point(aes(col=condition)) +
	geom_smooth(method="lm")+ theme_bw() +coord_equal()+ ggtitle("marker correlation") + facet_wrap(~condition) 
```

