---
title: "Predict missing marker"
author: "Attila Gabor"
date: "4/5/2019"
output: html_document
---
Here we build random forest model to predict the median p-ERK response to combination of 
stimulus and inhibitors. Predictions are based on other markers. 
The model is trained from cell-lines which contains all the markerks in many exper



```{r}
phospho_median <- readRDS("./data/raw_data/Median_allsamples_nocontrols_withcellcount.rds") %>% as_tibble()
reporters = c('IdU', 'p-CREB', 'p-STAT5', 'p-SRC', 'p-FAK', 'p-MEK', 'p-MAPKAPK2',
			  'p-S6K', 'p-MAP2K3', 'p-STAT1', 'p-p53', 'p-NFkB', 'p-p38', 'p-AMPK',
			  'p-Akt(Ser473)', 'p-ERK', 'CyclinB', 'p-GSK3b', 'GAPDH', 'p-MKK3-MKK6',
			  'p-PDPK1', 'p-BTK', 'p-p90RSK', 'p-SMAD23', 'p-STAT3', 'p-JNK',
			  'Ki-67', 'p-H3', 'p-S6', 'cleavedCas', 'p-MKK4', 'p-AKT(Thr308)',
			  'p-RB', 'p-4EBP1')
phospho_median <- phospho_median %>%
	rename(cellcount=`dmt$cellcount`) %>% 
	gather(key = "reporter",value="value", reporters) %>%
	mutate(reporter = make.names(reporter)) %>%
	mutate(time = ifelse(is.na(time),0,as.numeric(as.character(time)))) # %>%  # convert measured values to valid names to avoid later isssues 
#select(-cellcount) # we dont need this info
phospho_median
```

# collect here the cell_lines that must not be giving away:
for phospo-ERK:
- T47D
- HCC1806



## 1. Phospho-ERK prediction

```{r}
cell_lines = as.character(unique(phospho_median$cell_line))

set.seed(123)
validation_cell_lines = unique(c("T47D","HCC1806",sample(cell_lines,8)))
trainig_cell_lines = cell_lines[! cell_lines %in% validation_cell_lines]

# training data: <free to use>
# all reporters from training cell lines
pERK_training_data = phospho_median %>% 
	filter(cell_line %in% trainig_cell_lines) %>%
	spread(reporter,value)

# validation data: <KEEP>
# all reporters from validation cell lines
pERK_validation_data = phospho_median %>% 
	filter(cell_line %in% validation_cell_lines) %>%
	spread(reporter,value)

# training data: <use for predict validation data>
# all reporters, but p-ERK from validation cell lines
pERK_test_data = phospho_median %>% 
	filter(cell_line %in% validation_cell_lines) %>%
	spread(reporter,value) %>%
	select(-p.ERK)
```

## 1.1 RF simple psites
First we use only the p-Sites (apart from p-ERK). 
We dont use explicitly the time or treatment in this model. 

```{r}
library(ranger)

# we start from pERK_training_data

reporters <-  colnames(pERK_training_data)[6:39]

# predictors = c("treatment","time", reporters[reporters!="p.ERK"])
predictors = c(reporters[!reporters %in% c("p.ERK")])
response = "p.ERK"

rangerres0=ranger(as.formula(paste(response, " ~ ", paste(predictors, collapse= "+"))),
				  data = pERK_training_data, 
				  num.trees=1000,
				  importance = "impurity",
				  verbose=T,
				  num.threads = 3)

sort(rangerres0$variable.importance,decreasing = T)
```


This  shows that p-ERK is mainly determined by its downstream nodes


```{r, fig.height=7, fig.width=10}
pred.p.ERK_RF1 <- predict(rangerres0, data = pERK_validation_data)

pERK_validation_data %>% 
	select(cell_line,treatment,time,time_course,p.ERK) %>%
	bind_cols(p.ERK.pred = pred.p.ERK_RF1$predictions) %>%
	ggplot() +
	geom_point(aes(time,p.ERK,col="data")) + 
	geom_point(aes(time,p.ERK.pred,col="pred"))  + 
	geom_line(aes(time,p.ERK, col="data")) + 
	geom_line(aes(time,p.ERK.pred,col="pred")) +
	facet_grid(treatment~cell_line) +
	scale_color_manual(values = c("data" = "black", "pred" = "red")) + 
	theme_bw() + 
	ggtitle("p-ERK prediciton in validation data")
```

```{r, fig.width=9,fig.height=3}
pERK_validation_data %>% 
	select(cell_line,treatment,time,time_course,p.ERK) %>%
	bind_cols(p.ERK.pred = pred.p.ERK_RF1$predictions) %>%
	group_by(cell_line,treatment) %>%
	summarise(RMSE = sqrt(sum((p.ERK - p.ERK.pred)^2)/n()),
			  corr = cor(p.ERK, p.ERK.pred)) %>%
	ggplot() +
	geom_col(aes(x=treatment, y=RMSE,fill=treatment)) +
	geom_hline(aes(yintercept = mean(RMSE))) + 
	facet_grid(~cell_line)+ theme_bw()+ ylim(0,1) + 
	theme(axis.text.x = element_text(angle = 90, hjust = 1)) 
```


Model performance: the baseline model captures well the time-course of p-ERK. Bigger discrepancy can be seen at the iMEK condition. The cell-lines where pERK actually depends on p-MEK have the highest error. (this is bc drug that inhibits MEK does not reduce p-MEK. TODO inform RF by the DRUG)  


## 1.2 RF treatment information

Let's train the RF model with info on MEK inhibiton. 
The goal is that the RF can recognise that when we inhibit MEK, then p-MEK should 
not be used as the predictor of p-ERK. 

```{r}
# we start from pERK_training_data
RF2_data <- pERK_training_data %>% 
	mutate(p.MEK = ifelse(treatment=="iMEK",0,p.MEK))

reporters <-  colnames(RF2_data)[6:39]

#predictors = c("iMEK",reporters[!reporters %in% c("p.ERK")])
predictors = c(reporters[!reporters %in% c("p.ERK")])
response = "p.ERK"
#rangerres2=ranger(as.formula(paste(response, " ~ iMEK*p.MEK+ ", paste(predictors, collapse= "+"))),
rangerres2=ranger(as.formula(paste(response, " ~ ", paste(predictors, collapse= "+"))),
				  data = RF2_data, 
				  num.trees=1000,
				  importance = "impurity",
				  verbose=T,
				  num.threads = 3)

data.frame(importance = sort(rangerres2$variable.importance,decreasing = T))
```

```{r, fig.height=7, fig.width=10}
RF2_valid_data <- pERK_validation_data %>% mutate(p.MEK = ifelse(treatment=="iMEK",0,p.MEK))

pred.p.ERK_RF2 <- predict(rangerres2, data = RF2_valid_data)

RF2_valid_data %>% 
	select(cell_line,treatment,time,time_course,p.ERK) %>%
	bind_cols(p.ERK.pred = pred.p.ERK_RF2$predictions) %>%
	ggplot() +
	geom_point(aes(time,p.ERK,col="data")) + 
	geom_point(aes(time,p.ERK.pred,col="pred"))  + 
	geom_line(aes(time,p.ERK, col="data")) + 
	geom_line(aes(time,p.ERK.pred,col="pred")) +
	facet_grid(treatment~cell_line) +
	scale_color_manual(values = c("data" = "black", "pred" = "red")) + 
	theme_bw() + 
	ggtitle("p-ERK prediciton in validation data")

```

```{r,fig.width=9,fig.height=3}
RF2_valid_data %>% 
	select(cell_line,treatment,time,time_course,p.ERK) %>%
	bind_cols(p.ERK.pred = pred.p.ERK_RF2$predictions) %>%
	group_by(cell_line,treatment) %>%
	summarise(RMSE = sqrt(sum((p.ERK - p.ERK.pred)^2)/n()),
			  corr = cor(p.ERK, p.ERK.pred)) %>%
	ggplot() +
	geom_col(aes(x=treatment, y=RMSE,fill=treatment)) +
	geom_hline(aes(yintercept = mean(RMSE))) + 
	facet_grid(~cell_line)+ theme_bw()+ ylim(0,1)+
	theme(axis.text.x = element_text(angle = 90, hjust = 1)) 
```



## 1.3 RF treatment information, no downstream

Remove the nodes from the features that are downstream of ERK


```{r}
# we start from pERK_training_data
RF3_data <- pERK_training_data %>% 
	mutate(p.MEK = ifelse(treatment=="iMEK",0,p.MEK))

reporters <-  colnames(RF3_data)[6:39]

#predictors = c("iMEK",reporters[!reporters %in% c("p.ERK")])
predictors = c(reporters[!reporters %in% c("p.ERK", "p.p90RSK","p.MKK4","p.S6K","p.MAPKAPK2","p.p38")])
response = "p.ERK"
#rangerres2=ranger(as.formula(paste(response, " ~ iMEK*p.MEK+ ", paste(predictors, collapse= "+"))),
rangerres3=ranger(as.formula(paste(response, " ~ ", paste(predictors, collapse= "+"))),
				  data = RF3_data, 
				  num.trees=1000,
				  importance = "impurity",
				  verbose=T,
				  num.threads = 3)

data.frame(importance = sort(rangerres3$variable.importance,decreasing = T))
```

```{r, fig.height=7, fig.width=10}
RF3_valid_data <- pERK_validation_data %>% mutate(p.MEK = ifelse(treatment=="iMEK",0,p.MEK))

pred.p.ERK_RF3 <- predict(rangerres3, data = RF3_valid_data)

RF3_valid_data %>% 
	select(cell_line,treatment,time,time_course,p.ERK) %>%
	bind_cols(p.ERK.pred = pred.p.ERK_RF3$predictions) %>%
	ggplot() +
	geom_point(aes(time,p.ERK,col="data")) + 
	geom_point(aes(time,p.ERK.pred,col="pred"))  + 
	geom_line(aes(time,p.ERK, col="data")) + 
	geom_line(aes(time,p.ERK.pred,col="pred")) +
	facet_grid(treatment~cell_line) +
	scale_color_manual(values = c("data" = "black", "pred" = "red")) + 
	theme_bw() + 
	ggtitle("p-ERK prediciton in validation data")

```

```{r,fig.width=9,fig.height=3}
RF3_valid_data %>% 
	select(cell_line,treatment,time,time_course,p.ERK) %>%
	bind_cols(p.ERK.pred = pred.p.ERK_RF3$predictions) %>%
	group_by(cell_line,treatment) %>%
	summarise(RMSE = sqrt(sum((p.ERK - p.ERK.pred)^2)/n()),
			  corr = cor(p.ERK, p.ERK.pred)) %>% 
	ggplot() +
	geom_col(aes(x=treatment, y=RMSE,fill=treatment)) +
	geom_hline(aes(yintercept = mean(RMSE))) + 
	facet_grid(~cell_line)+ theme_bw()+ ylim(0,1.2)+
	theme(axis.text.x = element_text(angle = 90, hjust = 1)) + ggtitle(paste0("Prediction error"))
```





