---
title: "DREAM Challenge: supporting figures for EGF perturbation data"
author: "Attila Gabor and Marco Tognetti"
date: "3/20/2019"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```


## Overview : challenge 1.i

What is the input:

Challange 1.i is about predicting the time courses of some markers in some cell-lines based 
on other markers of the cell line. 

Cell-lines show various trends in response to stimuli and certain drug treatments. 
We categorised these trends as non-responsive, peaking, rising, plateau and decreasing.

In the following section we show some examples of these features from the data.

The task will be to learn the dependency among the time course markers and 
generate a model that can predict some missing markers. 




## Input data shaping: 
```{r, echo=FALSE}
library(pheatmap)
phospho_median <- readRDS("./data/raw_data/Median_allsamples_nocontrols_withcellcount.rds") %>% as_tibble()
reporters = c('IdU', 'p-CREB', 'p-STAT5', 'p-SRC', 'p-FAK', 'p-MEK', 'p-MAPKAPK2',
			  'p-S6K', 'p-MAP2K3', 'p-STAT1', 'p-p53', 'p-NFkB', 'p-p38', 'p-AMPK',
			  'p-Akt(Ser473)', 'p-ERK', 'CyclinB', 'p-GSK3b', 'GAPDH', 'p-MKK3-MKK6',
			  'p-PDPK1', 'p-BTK', 'p-p90RSK', 'p-SMAD23', 'p-STAT3', 'p-JNK',
			  'Ki-67', 'p-H3', 'p-S6', 'cleavedCas', 'p-MKK4', 'p-AKT(Thr308)',
			  'p-RB', 'p-4EBP1')
phospho_median <-phospho_median %>% rename(cellcount=`dmt$cellcount`)
phospho_median <-phospho_median %>% gather(key = "reporter",value="value",reporters) 
phospho_median <- phospho_median %>% 	mutate(time=ifelse(is.na(time),0,as.numeric(as.character(time))))
phospho_median

```

# Show Interesting features to support challenge questions

### Response to EGF
on absolute scale

```{r, fig.height=10, fig.width=10,eval=FALSE}
pdf("./figures/median_phospho_EGF_response.pdf")
for(marker in unique(phospho_median$reporter)){
	gg = phospho_median %>% 
	filter(treatment=="EGF", reporter == marker) %>%
	group_by(cell_line, time) %>% summarise(value=mean(value)) %>%
	ggplot(aes(time,value)) + 
	geom_line() +
	facet_wrap(~cell_line) + theme_bw() + guides(color=FALSE) + ggtitle(marker)
	print(gg)
}
dev.off()
```


### p-CREB
some cell-line show good activation/plateaue curves, others not reponding

```{r}
creb_responders = c("HCC1428", "HCC1569","HCC1599", "HCC2218","MDAMB468")
creb_nonresponders = c("MDAkb2","MACLS2", "T47D")
phospho_median %>% 
	filter(treatment=="EGF") %>%
	filter(reporter == "p-CREB") %>%
	filter(cell_line %in% c(creb_responders,creb_nonresponders)) %>%
	ggplot(aes(time,value,group=cell_line)) + 
	geom_line() +
	facet_wrap(~cell_line) + 
	theme_bw() + 
	guides(color=FALSE) +
	geom_smooth(formula = "y~x", method = "loess") +
	ggtitle("p-CREB",subtitle = "responding and non-responding cell-lines")
```

### p-STAT5
non-responsive or plateau effects, high range

```{r}
pSite = "p-STAT5"
responders = c("AU565", "HCC1806","HCC1187", "MCF12A","MDAMB468")
nonresponders = c("BT20","BT483", "T47D")
phospho_median %>% 
	filter(treatment=="EGF") %>%
	filter(reporter == pSite) %>%
	filter(cell_line %in% c(responders,nonresponders)) %>%
	ggplot(aes(time,value,group=cell_line)) + 
	geom_line() +
	facet_wrap(~cell_line) + 
	theme_bw() + 
	guides(color=FALSE) +
	geom_smooth(formula = "y~x", method = "loess") +
	ggtitle(pSite,subtitle = "responding and non-responding cell-lines")
```
### p-SRC
I dont see cell lines responding to EGF

### p-FAK
mostly non-responsive

```{r}
pSite = "p-FAK"
responders = c("DU4475", "MCF12A","BT20")
nonresponders = c("UACC893","KPL1", "HCC38")
phospho_median %>% 
	filter(treatment=="EGF") %>%
	filter(reporter == pSite) %>%
	filter(cell_line %in% c(responders,nonresponders)) %>%
	ggplot(aes(time,value,group=cell_line)) + 
	geom_line() +
	facet_wrap(~cell_line) + 
	theme_bw() + 
	guides(color=FALSE) +
	geom_smooth(formula = "y~x", method = "loess") +
	ggtitle(pSite,subtitle = "responding and non-responding cell-lines")
```
### p-MEK
bit noisy, small changes, maybe because of large range. 
```{r}
pSite = "p-MEK"
responders = c("AU565", "DU4475","HCC2157","MDAMB436")
nonresponders = c("BT483","KPL1", "MX1","ZR7530")
phospho_median %>% 
	filter(treatment=="EGF") %>%
	filter(reporter == pSite) %>%
	filter(cell_line %in% c(responders,nonresponders)) %>%
	ggplot(aes(time,value,group=cell_line)) + 
	geom_line() +
	facet_wrap(~cell_line) + 
	theme_bw() + 
	guides(color=FALSE) +
	geom_smooth(formula = "y~x", method = "loess") +
	ggtitle(pSite,subtitle = "responding and non-responding cell-lines")
```
### p-MAPKAPK2
various responses, not always clear. 
```{r}
pSite = "p-MAPKAPK2"
responders = c("HCC1187", "AU565","MCF10A","HCC1428")
nonresponders = c("MACLS2","KPL1", "ZR7530","MFM223")
phospho_median %>% 
	filter(treatment=="EGF") %>%
	filter(reporter == pSite) %>%
	filter(cell_line %in% c(responders,nonresponders)) %>%
	ggplot(aes(time,value,group=cell_line)) + 
	geom_line() +
	facet_wrap(~cell_line) + 
	theme_bw() + 
	guides(color=FALSE) +
	geom_smooth(formula = "y~x", method = "loess") +
	ggtitle(pSite,subtitle = "responding and non-responding cell-lines")
```


### p−S6K

```{r}
pSite = "p-S6K"
responders = c("AU565", "DU4475","BT483","HCC1187")
nonresponders = c("BT483","KPL1", "CAL120","MFM223","MPE600")
phospho_median %>% 
	filter(treatment=="EGF") %>%
	filter(reporter == pSite) %>%
	filter(cell_line %in% c(responders,nonresponders)) %>%
	ggplot(aes(time,value,group=cell_line)) + 
	geom_line() +
	facet_wrap(~cell_line) + 
	theme_bw() + 
	guides(color=FALSE) +
	geom_smooth(formula = "y~x", method = "loess") +
	ggtitle(pSite,subtitle = "responding and non-responding cell-lines")
```

### p−MAP2K3
pretty noisy and the range of values are narrow


### p−STAT1

```{r}
pSite = "p-STAT1"
responders = c("AU565", "HCC1806", "BT20","MDAMB468")
nonresponders = c("BT483", "T47D","HCC1187")
phospho_median %>% 
	filter(treatment=="EGF") %>%
	filter(reporter == pSite) %>%
	filter(cell_line %in% c(responders,nonresponders)) %>%
	ggplot(aes(time,value,group=cell_line)) + 
	geom_line() +
	facet_wrap(~cell_line) + 
	theme_bw() + 
	guides(color=FALSE) +
	geom_smooth(formula = "y~x", method = "loess") +
	ggtitle(pSite,subtitle = "responding and non-responding cell-lines")
```
## p−p53
does not have much dynamics
```{r}
pSite = "p-p53"
responders = c("184B5", "DU4475", "HCC1143","Hs578T")
nonresponders = c("BT483", "T47D","HCC1187")
phospho_median %>% 
	filter(treatment=="EGF") %>%
	filter(reporter == pSite) %>%
	filter(cell_line %in% c(responders,nonresponders)) %>%
	ggplot(aes(time,value,group=cell_line)) + 
	geom_line() +
	facet_wrap(~cell_line) + 
	theme_bw() + 
	guides(color=FALSE) +
	geom_smooth(formula = "y~x", method = "loess") +
	ggtitle(pSite,subtitle = "responding and non-responding cell-lines")
```

### p−NFkB

```{r}
pSite = "p-NFkB"
responders = c("184B5","AU565" ,"BT549", "CAL851","HCC1937","MCF12A","HCC1954")
nonresponders = c("BT483", "T47D","EFM19")
phospho_median %>% 
	filter(treatment=="EGF") %>%
	filter(reporter == pSite) %>%
	filter(cell_line %in% c(responders,nonresponders)) %>%
	ggplot(aes(time,value,group=cell_line)) + 
	geom_line() +
	facet_wrap(~cell_line) + 
	theme_bw() + 
	guides(color=FALSE) +
	geom_smooth(formula = "y~x", method = "loess") +
	ggtitle(pSite,subtitle = "responding and non-responding cell-lines")
```
### p−p38
we nice dynamics !!
```{r}
pSite = "p-p38"
responders = c("184B5","AU565" ,"BT549", "CAL851","HCC1937","MCF12A","HCC1954","BT483", "T47D")
nonresponders = c("EFM19")
phospho_median %>% 
	filter(treatment=="EGF") %>%
	filter(reporter == pSite) %>%
	filter(cell_line %in% c(responders,nonresponders)) %>%
	ggplot(aes(time,value,group=cell_line)) + 
	geom_line() +
	facet_wrap(~cell_line) + 
	theme_bw() + 
	guides(color=FALSE) +
	geom_smooth(formula = "y~x", method = "loess") +
	ggtitle(pSite,subtitle = "plateau and peaks")
```
### p−AMPK
increasing and decreaseing dynamics
```{r}
pSite = "p-AMPK"
increaseing = c("184A1","BT20",   "HCC1937", "HCC1937","HCC1937")
decreasing = c("CAL851", "HDQP1","MCF10F","ZR751","HDQP1")
phospho_median %>% 
	filter(treatment=="EGF") %>%
	filter(reporter == pSite) %>%
	filter(cell_line %in% c(increaseing,decreasing)) %>%
	ggplot(aes(time,value,group=cell_line)) + 
	geom_line() +
	facet_wrap(~cell_line) + 
	theme_bw() + 
	guides(color=FALSE) +
	geom_smooth(formula = "y~x", method = "loess") +
	ggtitle(pSite,subtitle = "up and downs")
```

### p-AKT_S473
nice signals as pointed out by Marco
```{r}
pSite = "p-Akt(Ser473)"
plateau = c("AU565","HBL100",   "HCC1937", "HCC70","Hs578T")
peaks = c("CAL851", "MCF12A","MDAkb2","UACC893","SKBR3")
phospho_median %>% 
	filter(treatment=="EGF") %>%
	filter(reporter == pSite) %>%
	filter(cell_line %in% c(plateau,peaks)) %>%
	ggplot(aes(time,value,group=cell_line)) + 
	geom_line() +
	facet_wrap(~cell_line) + 
	theme_bw() + 
	guides(color=FALSE) +
	geom_smooth(formula = "y~x", method = "loess") +
	ggtitle(pSite,subtitle = "up and downs")
```
### p-ERK
nice signaling 
```{r}
pSite = "p-ERK"
plateau = c("AU565",   "MDAMB436", "HCC1395","CAL51")
peaks = c("HBL100", "MCF12A","MDAkb2","UACC893","SKBR3")
phospho_median %>% 
	filter(treatment=="EGF") %>%
	filter(reporter == pSite) %>%
	filter(cell_line %in% c(plateau,peaks)) %>%
	ggplot(aes(time,value,group=cell_line)) + 
	geom_line() +
	facet_wrap(~cell_line) + 
	theme_bw() + 
	guides(color=FALSE) +
	geom_smooth(formula = "y~x", method = "loess") +
	ggtitle(pSite,subtitle = "up and downs")
```


## CyclinB
no much dynamics

### GSK3b  

```{r}
pSite = "p-GSK3b"
responders = c("184B5",   "CAL120", "AU565","MDAMB157","UACC893")
nonresponders = c("HBL100", "HDQP1","MDAkb2","T47D")
phospho_median %>% 
	filter(treatment=="EGF") %>%
	filter(reporter == pSite) %>%
	filter(cell_line %in% c(responders,nonresponders)) %>%
	ggplot(aes(time,value,group=cell_line)) + 
	geom_line() +
	facet_wrap(~cell_line) + 
	theme_bw() + 
	guides(color=FALSE) +
	geom_smooth(formula = "y~x", method = "loess") +
	ggtitle(pSite,subtitle = "responders and non-responsive nodes")
```

### GAPDH  
no strong response

### p-MKK3-MKK6
```{r}
pSite = "p-MKK3-MKK6"
responders = c("184A1", "HBL100",  "CAL120", "HCC2185","MDAMB157","HCC70")
nonresponders = c("AU565", "HDQP1","MDAkb2","T47D")
phospho_median %>% 
	filter(treatment=="EGF") %>%
	filter(reporter == pSite) %>%
	filter(cell_line %in% c(responders,nonresponders)) %>%
	ggplot(aes(time,value,group=cell_line)) + 
	geom_line() +
	facet_wrap(~cell_line) + 
	theme_bw() + 
	guides(color=FALSE) +
	geom_smooth(formula = "y~x", method = "loess") +
	ggtitle(pSite,subtitle = "responders and non-responsive nodes")
```
### p-PDPK1
effect is not really clear
```{r}
pSite = "p-PDPK1"
responders = c("184B5", "MX1")
nonresponders = c("HCC1806","T47D")
phospho_median %>% 
	filter(treatment=="EGF") %>%
	filter(reporter == pSite) %>%
	filter(cell_line %in% c(responders,nonresponders)) %>%
	ggplot(aes(time,value,group=cell_line)) + 
	geom_line() +
	facet_wrap(~cell_line) + 
	theme_bw() + 
	guides(color=FALSE) +
	geom_smooth(formula = "y~x", method = "loess") +
	ggtitle(pSite,subtitle = "responders and non-responsive nodes")
```
### p-BTK
effect is not really clear
```{r}
pSite = "p-BTK"
responders = c("HCC2185","DU4475", "HCC2185","MCF7","ZR7530","184A1")
nonresponders = c("MDAkb2","T47D")
phospho_median %>% 
	filter(treatment=="EGF") %>%
	filter(reporter == pSite) %>%
	filter(cell_line %in% c(responders,nonresponders)) %>%
	ggplot(aes(time,value,group=cell_line)) + 
	geom_line() +
	facet_wrap(~cell_line) + 
	theme_bw() + 
	guides(color=FALSE) +
	geom_smooth(formula = "y~x", method = "loess") +
	ggtitle(pSite,subtitle = "responders and non-responsive nodes")
```

### p-p90RSK
good signals
```{r}
pSite = "p-p90RSK"
plateau = c("AU565",   "MDAMB436", "HCC1395","CAL51")
peaks = c("HBL100", "MCF12A","MDAkb2","UACC893","SKBR3")
phospho_median %>% 
	filter(treatment=="EGF") %>%
	filter(reporter == pSite) %>%
	filter(cell_line %in% c(plateau,peaks)) %>%
	ggplot(aes(time,value,group=cell_line)) + 
	geom_line() +
	facet_wrap(~cell_line) + 
	theme_bw() + 
	guides(color=FALSE) +
	geom_smooth(formula = "y~x", method = "loess") +
	ggtitle(pSite,subtitle = "responders and non-responsive nodes")
```
### p-SMAD23
effect is not really clear

```{r}
pSite = "p-SMAD23"
samples = c("DU4475",   "CAL120", "HCC70","T47D")

phospho_median %>% 
	filter(treatment=="EGF") %>%
	filter(reporter == pSite) %>%
	filter(cell_line %in% c(samples)) %>%
	ggplot(aes(time,value,group=cell_line)) + 
	geom_line() +
	facet_wrap(~cell_line) + 
	theme_bw() + 
	guides(color=FALSE) +
	geom_smooth(formula = "y~x", method = "loess") +
	ggtitle(pSite,subtitle = "responders and non-responsive nodes")
```


### p-STAT3
clearly most of the time no response, sometimes good peaks. 

```{r}
pSite = "p-STAT3"
responders = c("184A1",   "184B5", "BT20","HCC1806","MDAMB468")
nonresponders = c("HCC2218","MPE600")
phospho_median %>% 
	filter(treatment=="EGF") %>%
	filter(reporter == pSite) %>%
	filter(cell_line %in% c(responders,nonresponders)) %>%
	ggplot(aes(time,value,group=cell_line)) + 
	geom_line() +
	facet_wrap(~cell_line) + 
	theme_bw() + 
	guides(color=FALSE) +
	geom_smooth(formula = "y~x", method = "loess") +
	ggtitle(pSite,subtitle = "responders and non-responsive nodes")
```


### p-JNK
effect is not really clear

```{r}
pSite = "p-JNK"
responders = c("184A1",   "184B5", "BT549","MCF10A","SKBR3")
nonresponders = c("HCC1806","MDAMB453")

phospho_median %>% 
	filter(treatment=="EGF") %>%
	filter(reporter == pSite) %>%
	filter(cell_line %in% c(responders,nonresponders)) %>%
	ggplot(aes(time,value,group=cell_line)) + 
	geom_line() +
	facet_wrap(~cell_line) + 
	theme_bw() + 
	guides(color=FALSE) +
	geom_smooth(formula = "y~x", method = "loess") +
	ggtitle(pSite,subtitle = "responders and non-responsive nodes")
```
### Ki-67
not much dynamics


### p-H3
effect is not really clear

```{r}
pSite = "p-H3"

responders = c("184A1",   "184B5", "HCC1937","MCF10A","SKBR3","UACC893")
nonresponders = c("BT549","T47D","DU4475")

phospho_median %>% 
	filter(treatment=="EGF") %>%
	filter(reporter == pSite) %>%
	filter(cell_line %in% c(responders,nonresponders)) %>%
	ggplot(aes(time,value,group=cell_line)) + 
	geom_line() +
	facet_wrap(~cell_line) + 
	theme_bw() + 
	guides(color=FALSE) +
	geom_smooth(formula = "y~x", method = "loess") +
	ggtitle(pSite,subtitle = "responders and non-responsive nodes")
```
### p-S6
really nice signals

```{r}
pSite = "p-S6"

responders = c("184A1",   "HCC1187", "MCF12A","MCF10A","SKBR3","UACC893")
nonresponders = c("CAL51","EFM192A","MACLS2")

phospho_median %>% 
	filter(treatment=="EGF") %>%
	filter(reporter == pSite) %>%
	filter(cell_line %in% c(responders,nonresponders)) %>%
	ggplot(aes(time,value,group=cell_line)) + 
	geom_line() +
	facet_wrap(~cell_line) + 
	theme_bw() + 
	guides(color=FALSE) +
	geom_smooth(formula = "y~x", method = "loess") +
	ggtitle(pSite,subtitle = "responders and non-responsive nodes")
```

### cleavedCas
not clear.

### p-MKK4
nice signals ! plateau and peaks 


```{r}
pSite = "p-MKK4"

responders = c("AU565",   "HCC2218", "CAL120","HCC1187","T47D","MDAMB157")
nonresponders = c("MCF7","HCC1500","HCC70")

phospho_median %>% 
	filter(treatment=="EGF") %>%
	filter(reporter == pSite) %>%
	filter(cell_line %in% c(responders,nonresponders)) %>%
	ggplot(aes(time,value,group=cell_line)) + 
	geom_line() +
	facet_wrap(~cell_line) + 
	theme_bw() + 
	guides(color=FALSE) +
	geom_smooth(formula = "y~x", method = "loess") +
	ggtitle(pSite,subtitle = "responders and non-responsive nodes")
```


### p−AKT(Thr308)
many unclear but some are nice .


```{r}
pSite = "p-AKT(Thr308)"

responders = c("MCF12A",   "HCC1143", "DU4475","Hs578T","T47D")
nonresponders = c("MCF7","HCC1500","HCC70")

phospho_median %>% 
	filter(treatment=="EGF") %>%
	filter(reporter == pSite) %>%
	filter(cell_line %in% c(responders,nonresponders)) %>%
	ggplot(aes(time,value,group=cell_line)) + 
	geom_line() +
	facet_wrap(~cell_line) + 
	theme_bw() + 
	guides(color=FALSE) +
	geom_smooth(formula = "y~x", method = "loess") +
	ggtitle(pSite,subtitle = "responders and non-responsive nodes")
```

### 4EBP1 and RB
not much dynamics
