---
title: "Exploring proteomics for DREAM"
author: "Attila Gabor"
date: "4/29/2019"
output:
  html_document: 
    toc: yes
  pdf_document: default
---
Checking the proteomics data for the DREAM challenge. 

```{r,include=FALSE}
library(tidyverse)
library(ggplot2)
```

```{r}
proteomics_raw <- readRDS("data/proteomics/MSstat_groupComparison_selceted.rds") %>% as_tibble()
proteomics_raw
```

```{r}
str(proteomics_raw)
unique(proteomics_raw$Label)
```

Columns:
- Protein: uniprotIDs divided by ";". 
	- Why are multiple proteins in the same row?
	- in total 9031 protein families measured -- how does it distribute across cell-lines?
- Label: cell-line info


How many proteins measured per cell-line?
```{r}
protein_cell_line_table <- table(proteomics_raw[,1:2])
all(protein_cell_line_table==1) == TRUE
```
apparently all the proteins are measured across all cell-lines? -- appears in the table, maybe some values are NA/Inf.

```{r}
print(nrow(proteomics_raw))
print(table(proteomics_raw$issue))
```
Print a heatmap to show missing elements:
```{r,fig.width=10}
measured_conds <- proteomics_raw[,1:2]
measured_conds <- measured_conds[is.na(proteomics_raw$issue),]
real_proteomics_data <- table(measured_conds)
pheatmap::pheatmap(real_proteomics_data,cluster_cols = F,cluster_rows = F,labels_row = "")
```
Some proteins seems to be missing from the majority of the cell-lines:
```{r}
rowSums(real_proteomics_data) %>% enframe() %>% arrange(value) %>% mutate(name=factor(name,levels = .$name)) %>%
	ggplot() + geom_point(aes(1:length(name),value)) + 
	xlab("protein rank") + ylab("measured in cell-lines") +
	ggtitle("Measured proteins without an issue",subtitle = "in how many cell-lines a protein was measured")
```
```{r}
any(proteomics_raw$ImputationPercentage>0)
```

