---
title: "Leaderboards for Subchallenge I"
author: "Attila Gabor"
date: "11/25/2019"
output: BiocStyle::html_document
---

```{r setup, include=FALSE}

library(RColorBrewer)
library(tidyverse)
library(ggplot2)
library(ggrepel)

source("../submission_analysis/utilities.R")

colors  = brewer.pal(n = 8,"Dark2")

```



# Overview
Here we report some statistics about subchallenge I and evaluate the leaderboard. 

```{r message=FALSE, warning=FALSE, paged.print=FALSE, results=FALSE, echo=FALSE}

LB_sc1 <- read_csv("../submission_data/final_round/outdated/leaderboard_all_rounds_sc1.csv") %>%
	fix_leaderboard_raw() %>% filter(!is.na(round)) %>%
	mutate(createdOn = fix_submission_dates(createdOn)) %>%
	mutate(round = paste0("round_",round))
LB_sc2 <- read_csv("../submission_data/final_round/outdated/leaderboard_all_rounds_sc2.csv") %>%
	fix_leaderboard_raw() %>%  filter(!is.na(round)) %>%
	mutate(createdOn = fix_submission_dates(createdOn)) %>%
	mutate(round = paste0("round_",round))
LB_sc3 <- read_csv("../submission_data/final_round/outdated/leaderboard_all_rounds_sc3.csv") %>%
	fix_leaderboard_raw() %>%  filter(!is.na(round)) %>%
	mutate(createdOn = fix_submission_dates(createdOn)) %>%
	mutate(round = paste0("round_",round))
LB_sc4 <- read_csv("../submission_data/final_round/outdated/leaderboard_all_rounds_sc4.csv") %>%
	fix_leaderboard_raw() %>%  filter(!is.na(round)) %>%
	mutate(createdOn = fix_submission_dates(createdOn)) %>%
	mutate(round = paste0("round_",round))
```

Some statistics about challenge I: 
```{r, echo=FALSE}
# Statistics: 
n_submissions <- LB_sc1 %>% filter(status== "SCORED") %>% nrow()
#print(paste("# Total Submissions:",n_submissions))

n_submission_by_round <- LB_sc1 %>% filter(status== "SCORED") %>% pull(round) %>% table()
#print(paste("# Total Submissions per round:"))
#print(n_submission_by_round)

n_teams <- LB_sc1 %>% filter(status== "SCORED") %>% pull(submitterId)%>% unique() %>% length()
#print(paste("# Teams:",n_teams))
```
- the number of total submissions: `r n_submissions`. 
- the number of submissions per rounds
	- round 1: `r n_submission_by_round[[1]]`
	- round 2: `r n_submission_by_round[[2]]`
	- round 3: `r n_submission_by_round[[3]]`
- the number of teams participated in this sub challenge: `r n_teams`

Submissions that are better then the reference models improved from round 1 to round 2, but not to final round. 

```{r, echo=FALSE, warning=FALSE }
### Subchallange I -------------------------------------------------------------
#1.1 score distribution across rounds
LB_sc1 %>% filter(status=="SCORED") %>% ggplot() + 
	geom_rect(xmin=0.7, xmax=1.25, ymin=0, ymax=40, fill ="#F8F8CE",color="black")+
	geom_histogram(aes(score),bins = 50, fill = colors[[1]]) + theme_bw() + 
	ggtitle("Sub-challenge I: predict missing marker") +
	ylab("# of submissions") + 
	geom_vline(xintercept = 0.903103,color = colors[[2]],size=2) + 
	geom_vline(xintercept = 3.992752,color = colors[[3]],size=2) + coord_cartesian(xlim = c(0,6)) + facet_grid( round~ .)

LB_sc1 %>% filter(score<1.2) %>% ggplot() + geom_histogram(aes(score),bins = 40, fill = colors[[1]]) + theme_bw() + 
	ggtitle("Sub-challenge I: predict missing marker") +
	ylab("# of submissions") + 
	geom_vline(xintercept = 0.903103,color = colors[[2]],size = 2) + 
	xlim(0.7,1.25) + theme(panel.background = element_rect(fill = "#F8F8CE" ))  + facet_grid( round~ .)

```



# Leaderboards


Next figure shows all the submissions. We report the scores of each team, 
the y-axis is capped at 1.2 for better visualisation around the best teams. 
We ordered the teams by their last submission as we agreed to rank them accordingly. 
The shape of the points indicate in which round they submitted that predictions. 

Some teams did not submit in the final rounds in some cases the final submissions were
worst than the previous ones. 

Here we go though the best teams as they are ranked on the figure. In all cases the write-ups need to be checked. 

- The first place is clear
- 2nd and 3rd positions: `NAD` and `@duct` is the same (NAD is the team, @duct is a member of the team)
- `ajay_scic` and `Sleeping` scored very poorly in the final round
- `Rhagave_india_SCS` didnt participate in the final round, but submitted write-up

TODO: decide what to do with the teams that didnt submit in the final round. This 
could influence the 3rd place. 


```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
LB_sc1 %>% filter(status == "SCORED") %>% 
	group_by(submitterId) %>%
	mutate(latest = createdOn == createdOn[which.max(createdOn)] ) %>%
	ungroup() %>%
	mutate(submitterId = factor(submitterId,levels = filter(.,latest) %>%arrange(score) %>% pull(submitterId)))%>%
	ggplot(aes(submitterId,score)) + 
	geom_point(aes(color=latest,shape=round),size = 2) + 
	theme_bw() + 
	theme(axis.text.x = element_text(angle = 90,hjust = 1)) +
	ggtitle("Submissions ordered by the latest score")

LB_sc1 %>% filter(status == "SCORED") %>% 
	group_by(submitterId) %>%
	mutate(latest = createdOn == createdOn[which.max(createdOn)] ) %>%
	ungroup() %>%
	mutate(submitterId = factor(submitterId,levels = filter(.,latest) %>%arrange(score) %>% pull(submitterId)))%>%
	ggplot(aes(submitterId,score)) + 
	geom_point(aes(color=latest,shape=round),size = 2) + 
	theme_bw() + 
	theme(axis.text.x = element_text(angle = 90,hjust = 1)) +
	ggtitle("Submissions ordered by the latest score (scores < 1.2)") + coord_cartesian(ylim = c(0.8,1.2))
```

## Strict leaderboard

- only teams with submission to round 3
- judging by their latest submission. 

```{r paged.print=FALSE, echo=FALSE}
LB_sc1 %>% filter(status == "SCORED",round == "round_3") %>% 
	group_by(submitterId) %>%
	mutate(latest = createdOn == createdOn[which.max(createdOn)] ) %>%
	ungroup() %>%
	mutate(submitterId = factor(submitterId,levels = filter(.,latest) %>%arrange(score) %>% pull(submitterId)))%>%
	ggplot(aes(submitterId,score)) + 
	geom_point(aes(color=latest,shape=round),size = 2) + 
	theme_bw() + 
	theme(axis.text.x = element_text(angle = 90,hjust = 1)) +
	ggtitle("Submissions ordered by the latest score (scores < 1.2)") + coord_cartesian(ylim = c(0.8,1.2))


LB_sc1 %>% filter(status == "SCORED",round == "round_3") %>% 
	group_by(submitterId) %>%
	mutate(latest = createdOn == createdOn[which.max(createdOn)] ) %>%
	ungroup() %>%
	filter(latest == TRUE) %>%
	arrange(score) %>%
	select(-objectId, -status, -FAILURE_REASON, -round,-latest) %>% print(.,n=23)
```

## Leaderboard based on best score

- teams with any submission 
- judging by the best score.


```{r paged.print=FALSE, echo=FALSE}

LB_sc1 %>% filter(status == "SCORED") %>% 
	group_by(submitterId) %>%
	top_n(-1, score) %>%
	filter(!duplicated(submitterId)) %>%  # for duplicates
	ungroup() %>%
	arrange(score) %>%
	mutate(submitterId = factor(submitterId,levels = arrange(.,score) %>% pull(submitterId)))%>%
	ggplot(aes(submitterId,score)) + 
	geom_point(aes(color=round, shape=round),size = 2) + 
	theme_bw() + 
	theme(axis.text.x = element_text(angle = 90,hjust = 1)) +
	ggtitle("Submissions ordered by the best scores (scores < 1.2)") +
	coord_cartesian(ylim = c(0.8,1.2))


LB_sc1 %>% filter(status == "SCORED") %>% 
	group_by(submitterId) %>%
	top_n(-1, score) %>%
	filter(!duplicated(submitterId)) %>%  # for duplicates
	ungroup() %>%
	arrange(score) %>%
	select(-objectId, -status, -FAILURE_REASON) %>%
	print(.,n=42)
```


## Leaderboard based on latest submission


- teams with any submission 
- judging by the latest score.


```{r paged.print=FALSE, echo=FALSE}

LB_sc1 %>% filter(status == "SCORED") %>% 
	group_by(submitterId) %>%
	mutate(latest = createdOn == createdOn[which.max(createdOn)] ) %>%
	filter(latest) %>%
	ungroup() %>%
	arrange(score) %>%
	mutate(submitterId = factor(submitterId,levels = arrange(.,score) %>% pull(submitterId)))%>%
	ggplot(aes(submitterId,score)) + 
	geom_point(aes(color=round, shape=round),size = 2) + 
	theme_bw() + 
	theme(axis.text.x = element_text(angle = 90,hjust = 1)) +
	ggtitle("Submissions ordered by the best scores (scores < 1.2)") +
	coord_cartesian(ylim = c(0.8,1.2))


LB_sc1 %>% filter(status == "SCORED") %>% 
	group_by(submitterId) %>%
	mutate(latest = createdOn == createdOn[which.max(createdOn)] ) %>%
	filter(latest) %>% 
	ungroup() %>%
	arrange(score) %>%
	select(-objectId, -status, -FAILURE_REASON) %>%
	print(.,n=42)
```




