---
title: "Leaderboards for Subchallenge IV"
author: "Attila Gabor"
date: "11/25/2019"
output: BiocStyle::html_document
---

```{r setup, include=FALSE}

library(RColorBrewer)
library(tidyverse)
library(ggplot2)
library(ggrepel)

source("../submission_analysis/utilities.R")

colors  = brewer.pal(n = 8,"Dark2")

```

# Overview
Here we report some statistics about subchallenge IV and evaluate the leaderboard. 

```{r message=FALSE, warning=FALSE, paged.print=FALSE, results=FALSE, echo=FALSE}
LB <- read_csv("../submission_data/final_round/outdated/leaderboard_all_rounds_sc4.csv") %>%
	fix_leaderboard_raw() %>%  filter(!is.na(round)) %>%
	mutate(createdOn = fix_submission_dates(createdOn)) %>%
	mutate(round = paste0("round_",round))
```

Some statistics about challenge IV: 
```{r, echo=FALSE}
# Statistics: 
n_submissions <- LB %>% filter(status== "SCORED") %>% nrow()
n_submission_by_round <- LB %>% filter(status== "SCORED") %>% pull(round) %>% table()
n_teams <- LB %>% filter(status== "SCORED") %>% pull(submitterId)%>% unique() %>% length()
```

- the number of total submissions: `r n_submissions`. 
- the number of submissions per rounds
	- round 1: `r n_submission_by_round[[1]]`
	- round 2: `r n_submission_by_round[[2]]`
	- round 3: `r n_submission_by_round[[3]]`
- the number of teams participated in this sub challenge: `r n_teams`

Nice improvement from roun 1 to round 2 and final round. 

```{r, echo=FALSE, warning=FALSE }
#1.1 score distribution across rounds
LB %>% filter(status=="SCORED") %>%
	ggplot() + 
	geom_histogram(aes(score),bins = 50, fill = colors[[1]]) + theme_bw() + 
	ggtitle("Sub-challenge IV: predict dynamic response from static") +
	ylab("# of submissions") + 
	geom_vline(xintercept = 0.340827,color = colors[[2]],size=2) + 
	geom_vline(xintercept = 2.95,color = colors[[3]],size=2) + facet_grid( round~ .)

```



# Leaderboards


Next figure shows all the submissions. We report the scores of each team, ordered
by their last submission as we agreed to rank them. 
The shape of the points indicate in which round they submitted that predictions. 

Some teams did not submit in the final rounds in some cases the final submissions were
worst than the previous ones. 

Here we go though the best teams as they are ranked on the figure. In all cases the write-ups need to be checked. 

- The top 3 places are stable. 
- `dreamskys` has a good prediction in round 3 that would be place 4, but not the latest submission.  


TODO: decide what to do with the teams that didnt submit in the final round. This 
could influence the 3rd place. 


```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
LB %>% filter(status == "SCORED") %>% 
	group_by(submitterId) %>%
	mutate(latest = createdOn == createdOn[which.max(createdOn)] ) %>%
	ungroup() %>%
	mutate(submitterId = factor(submitterId,levels = filter(.,latest) %>% arrange(score) %>% pull(submitterId)))%>%
	ggplot(aes(submitterId,score)) + 
	geom_point(aes(color=latest,shape=round),size = 2) + 
	theme_bw() + 
	theme(axis.text.x = element_text(angle = 90,hjust = 1)) +
	ggtitle("Submissions ordered by the latest score")

```

## Strict leaderboard

- only teams with submission to round 3
- judging by their latest submission. 

```{r paged.print=FALSE, echo=FALSE}
LB %>% filter(status == "SCORED",round == "round_3") %>% 
	group_by(submitterId) %>%
	mutate(latest = createdOn == createdOn[which.max(createdOn)] ) %>%
	ungroup() %>%
	mutate(submitterId = factor(submitterId,levels = filter(.,latest) %>%arrange(score) %>% pull(submitterId)))%>%
	ggplot(aes(submitterId,score)) + 
	geom_point(aes(color=latest,shape=round),size = 2) + 
	theme_bw() + 
	theme(axis.text.x = element_text(angle = 90,hjust = 1)) +
	ggtitle("Submissions ordered by the latest score")


LB %>% filter(status == "SCORED",round == "round_3") %>% 
	group_by(submitterId) %>%
	mutate(latest = createdOn == createdOn[which.max(createdOn)] ) %>%
	ungroup() %>%
	filter(latest == TRUE) %>%
	arrange(score) %>%
	select(-objectId, -status, -FAILURE_REASON, -round,-latest) %>% print(.,n=15)
```

## Leaderboard based on best score

- teams with any submission 
- judging by the best score.


```{r paged.print=FALSE, echo=FALSE}

LB %>% filter(status == "SCORED") %>% 
	group_by(submitterId) %>%
	top_n(-1, score) %>%
	filter(!duplicated(submitterId)) %>%  # for duplicates
	ungroup() %>%
	arrange(score) %>%
	mutate(submitterId = factor(submitterId,levels = arrange(.,score) %>% pull(submitterId)))%>%
	ggplot(aes(submitterId,score)) + 
	geom_point(aes(color=round, shape=round),size = 2) + 
	theme_bw() + 
	theme(axis.text.x = element_text(angle = 90,hjust = 1)) +
	ggtitle("Submissions ordered by the best scores") 


LB %>% filter(status == "SCORED") %>% 
	group_by(submitterId) %>%
	top_n(-1, score) %>%
	filter(!duplicated(submitterId)) %>%  # for duplicates
	ungroup() %>%
	arrange(score) %>%
	select(-objectId, -status, -FAILURE_REASON) %>%
	print(.,n=25)
```


## Leaderboard based on latest submission


- teams with any submission 
- judging by the latest score.


```{r paged.print=FALSE, echo=FALSE}

LB %>% filter(status == "SCORED") %>% 
	group_by(submitterId) %>%
	mutate(latest = createdOn == createdOn[which.max(createdOn)] ) %>%
	filter(latest) %>%
	ungroup() %>%
	arrange(score) %>%
	mutate(submitterId = factor(submitterId,levels = arrange(.,score) %>% pull(submitterId)))%>%
	ggplot(aes(submitterId,score)) + 
	geom_point(aes(color=round, shape=round),size = 2) + 
	theme_bw() + 
	theme(axis.text.x = element_text(angle = 90,hjust = 1)) +
	ggtitle("Submissions ordered by the best scores")


LB %>% filter(status == "SCORED") %>% 
	group_by(submitterId) %>%
	mutate(latest = createdOn == createdOn[which.max(createdOn)] ) %>%
	filter(latest) %>% 
	ungroup() %>%
	arrange(score) %>%
	select(-objectId, -status, -FAILURE_REASON) %>%
	print(.,n=42)
```




